\section{\label{sec:streamAPI}The \streams Core Classes}
The \streams framework provides a wide range of implementations for
data streams and processors. These are useful for reading application
data and defining a complete data flow.

In this section we provide a comprehensive overview of the classes and
implementations already available in the \streams library. These can
directly be used to design stream processes for various application
domains.

\input{stream_io_index}

\subsection{The {\em stream-core} Processor Classes}

The core packages of the \streams framework provide a set of processor
implementations that cover a lot of general stream processing
tasks. These processors serve as basic building blocks to design a
stream process. A processor can simply be added to a process by adding
an XML element with the name of the processor to the process element
as shown in Figure \ref{fig:exampleProcessor}.

\begin{figure}[h!]
\centering
\begin{lstlisting}{lang=xml}
     <process input="my-stream">
         <!--  convert the string of attribute x1 to a double value -->
         <stream.parser.ParseDouble key="x1" />
     </process>
\end{lstlisting}
\caption{\label{fig:exampleProcessor}The processor {\ttfamily stream.parser.ParseDouble} added to a process.}
\end{figure}

Based on their purpose, the processors are organized into packages. To
shorten the XML declaration, several packages are automatically
checked for when resolve a processor. For example the {\ttfamily stream.parser} package is among the default base packages. This allows for leaving out the package name, when adding the processor. Thus, the XML from Figure \ref{fig:exampleProcessor} is equivalent to the following XML snippet:
\begin{figure}[h!]
\centering
\begin{lstlisting}{lang=xml}
     <process input="my-stream">
         <!--  convert the string of attribute x1 to a double value -->
         <ParseDouble key="x1" />
     </process>
\end{lstlisting}
\end{figure}

The default packages that are automatically checked for when resolving processor names are:
\begin{itemize}
  \item {\ttfamily stream.data}
  \item {\ttfamily stream.flow}
  \item {\ttfamily stream.parser}
  \item {\ttfamily stream.script}.
\end{itemize}

\input{stream_flow_index}

\input{stream_data_index}

\input{stream_parser_index}

\input{stream_script_index}

\input{stream_image_index}