\clearpage
\section{\label{sec:machineLearning}Machine Learning with Continuous Data}
As the large volumes of data are merely handable with automatic
processing of that data, they are far away from being inspected
manually. On the other hand gaining insight from that data is the key
problem in various application domains.

Machine learning and data mining has put forth a plethora of
techniques and algorithms for pattern recognition, learning of
prediction model or clustering that all aim at exactly that key
problem: knowledge discover from data.

For the setting of continuous data, various algorithms have been
proposed which solve basic tasks inherent to the knowledge discovery
process as well as complex methods that allow for training classifiers
or finding clusters on steady streams of data. In this section we will
give an overview of the machine learning algorithms that have been
implemented within the \streams framework as well as the integration
of an existing library, namely the MOA \cite{moa}, which provides a
rich set of state of the art online learning algorithms. 

After that overview in Section \ref{sec:onlineLearning}, we discuss
the problem of the {\em online application} of machine learning models
which have either been trained online of offline, and then are used to
make real-life predictions on streaming data. This will be covered in
Section \ref{sec:onlineApplication}.

\subsubsection*{Notation used}
Within this section, we will denote each data item $d_i$ obtained from
a stream as a tuple $d_i = (d_{i,1},\ldots,d_{i,k}) \subset
M_1\times\ldots\times M_k$, where each $M_i$ might refer to some
domain, e.g. $M_i \subseteq \mathbb{R}$ or an arbitrary set. Further we will refer to the index $i$ as the index of the item $d_i$ in a data stream $D$, i.e.
\begin{displaymath}
  D = \langle d_0,d_1,\ldots \rangle.
\end{displaymath}


%%
%%
%%
\subsection{\label{sec:onlineLearning}Online Learning from Data Streams}
Learning from unbounded and continuous data poses tight challenges to
the designer of machine learning algorithms. Even simple basic
building blocks like the computation of a median or minima/maxima
values that might be required in a learning algorithms tend to become
difficult.


\subsubsection*{Approximating Distributions}
As a simple example, the {\em NaiveBayes}\cite{NB} classifier often
offers a adequate prediction performance in a lot of application
domains. Starting with the independence assumption of attributes, it
maximizes computes the class probabilities given the observed
attributes of a set of training instances. The base rule of bayes is
given in equation (\ref{eqn:naiveBayes}):
\begin{eqnarray}
  P(c | f_1,\ldots,f_n ) = \frac{P(c)\cdot P(f_1,\ldots,f_n|c)}{P(f_1,\ldots,f_n)}.\label{eqn:naiveBayes}
\end{eqnarray}

Assuming a fixed set $C = {c_1,\ldots,c_l}$ of observable classes, we
can easily approximate $P(c)$ by counting the occurences for each
$c_i$ in the observed stream. For pure numerical attributes $f_i$,
generally a gaussian normal distribution is assumed, such that the
factors $P(f_1,\ldots,f_n|c)$ and $P(f_1,\ldots,f_n)$ can be derived
by estimating the mean and average for each attribute $f_i$.

The settings gets a bit more interesting if the $f_i$ are nominal
values, such as variable strings from an unbounded domain such as
URLs, i.e. $M_i \cong \mathbb{N}$. In this case we cannot simply
derive a probability for each instance of an attribute as this would
require counting an unbounded set of strings, which clearly violates
the stream processing contraints mentioned in Section
\ref{sec:streamSetting}.

%%
%% How are classifiers embedded into the streams framework?
%% How can they be used?
%%
\subsubsection{Embedding Classifiers in \streams}


%%
%% Which classifiers/clusterers/etc. are available?
%%
\subsubsection{Available Online Learning Algorithms}

\subsubsection*{Online Statistics (Counting, Quantiles)}

\subsubsection*{Online Classifier}




%%
%% What is online model application? When is it required?
%% What methods are provided? RapidMiner-Models?
%%
\subsection{\label{sec:onlineApplication}Model Application on Data Streams}

%%
%% Raise some questions on "why" and "when" online model application
%% is required, touch "concept drift" problems,...
%%
%% keywords: monitoring, static models, expert knowledge?
%%

%%
%% A simple but important use of online model application is the 
%% evaluation of classifiers on online data streams.
%%
%% This section should give an example walk-through for test-then-train
%% and demonstrate an example container that evaluates a classifier (MOA?)
%% on a data stream.
%%
\subsubsection{Evaluating Classifiers on a Stream}




%%
%% Section about the need to create a model on offline data and
%% using that model on dynamic continuous data streams
%%
\subsubsection{Learning Offline, Predicting Online}


