%\documentclass[12pt,parskip=half]{scrartcl}
\documentclass{TechReport}

% Select Sans serif font "Computer Modern Bright"
%\usepackage[T1]{fontenc}
%\usepackage{cmbright}

% ***TODO*** Select the right encoding for your document
%\usepackage[utf8]{inputenc}
%\usepackage[latin1]{inputenc}

%\usepackage{graphicx}

% Bibtex citation package
%\usepackage{cite}

%\usepackage{url}

% No page numbers here; these will be added for final Tech Report with
% all contributions
%\renewcommand*{\titlepagestyle}{empty}
%\pagestyle{empty}

%\title{LaTeX-Template for a Technical Report}
%\author{Christian Bockermann}
\date{01/2012}  % Serial Number/Year
\sfbproject{C3} % Your project identifier (two digit version)

\usepackage{amssymb}


% ***TODO*** Choose correct title, author, department and email address
\title{The \textsf{streams} Library for Processing Streaming Data}
\author{Christian Bockermann\\
  Lehrstuhl f\"ur k\"unstliche Intelligenz\\
  Technische Universit\"at Dortmund\\*[1ex]
  christian.bockermann@udo.edu}
% Please keep date empty
\date{}





\usepackage{listings}
\usepackage{hyperref}
\hypersetup{
  linktoc=all
}

\usepackage{tikz}
\usetikzlibrary{shapes} 
\usetikzlibrary{decorations,arrows} 
\usetikzlibrary{decorations.pathmorphing} 
\usepgflibrary{decorations.pathreplacing} 

\usepackage{algorithm}
%\usepackage{algorithmicx}
\usepackage{algpseudocode}

\usepackage{color}
\definecolor{rapidi}{RGB}{240,176,0}
\definecolor{rapidiText}{RGB}{102,51,0}
\definecolor{darkGreen}{RGB}{67,101,0}

\DeclareRobustCommand{\pointer}[1]{\tikz\node[draw,circle,draw=darkGreen,thick,fill=darkGreen!70,inner sep=1pt] at (0,-0.1) {\scriptsize\color{white}\textsf{#1}};} 

\lstset{language=XML,
%  otherkeywords={stream,container,process,service},
  basewidth={0.5em,0.45em},
  fontadjust=false,
  basicstyle=\footnotesize\ttfamily,
  keywordstyle=\color{blue},          % keyword style
  commentstyle=\ttfamily\color{black!60},       % comment style
  stringstyle=\color{darkGreen}
}


\newcommand{\todo}[1]{\marginpar{{\bf TODO:}\\#1}}
%\newcommand{\streams}{{\footnotesize\textsf{streams}\ }}
\newcommand{\streams}{{\em streams}\ }
\newcommand{\rapidminer}{{\footnotesize\textsf{RapidMiner}\ }}
%\newcommand{\plugin}{{\footnotesize\textsf{Streams Plugin}}}
\newcommand{\plugin}{{\em Streams Plugin}}

\newcommand{\defitem}[1]{\item[{\footnotesize\textbf{\textsf{#1}}}]}
\newcommand{\bigO}{\cO}
\newcommand{\barw}{\bar{w}}
\newcommand{\barbarw}{\bar w_M^\circ}
\newcommand{\vx}{x}


\begin{document}

\makesfbtitlepage

\newpage
\tableofcontents
\newpage

\begin{abstract}
{\normalsize
  In this report, we present the \streams library, a generic
  Java-based library for designing data stream processes. The \streams
  library defines a simple abstraction layer for data processing and
  provides a small set of online algorithms for counting and
  classification. Moreover it integrates existing libraries such as
  MOA. Processes are defined in XML files which follow the
  semantics and ideas of well established tools like Maven or the
  Spring Framework.

  The \streams library can be easily embedded into existing software,
  used as a standalone tool or be integrated into the RapidMiner tool
  using the \plugin.}
\end{abstract}

%\begin{document}
%\maketitle

%\input{introduction}

\section{Introduction\label{sec:intro}}
Todays applications produce a plethora of data on a continuous basis:
In the field of astrophysics, telescopes produce raw data that quickly
exceeds hundreds of gigabytes in a single day \cite{factTelescope}. In
the area of social networks, Twitter announced recently to have
observed a peak of 150,000 messages per minute during the TV duell
between president Obama and his challenger Romney \cite{}. In
computing farms and large scale networks there is a large pile of log
data about system state, system activities or working load gathered
and stored that is far from being analyzed by any human, but may
embody valuable information about the system
\cite{googleHDprediction}.

The data we are considering here, is non-stationary, in the sense that
new data is continously produced. This is fundamentally different to
the traditional batch-data processing paradigm that has been
dominating the data-analysis and data-mining world for the past
decades. That paradigm shift has triggered the evolving area of stream
or online analysis, which investigates algorithms for analysis of
continuous data.

Principles in DSM

In order to support the design of processes that allow for online
analysis we propose the \streams library. It is a Java based
abstraction layer for defining stream processes by means of simple XML
definitions. It generally follows the {\em pipes-and-filters} pattern
for data flow processing and extends this by an orthogonal control
flow view to support the notion of anytime services within the
process design.




%\input{intro}
%
%\input{relatedwork}

%\input{streamapi}

\input{process-design}

%\input{distributed}
%
%\input{streamslibrary}

%\input{stream-plugin}

\input{examples}
%\input{usecases}
%\input{appendix}

%\input{outlook}


% Bibliography using Bibtex, style plain
\bibliographystyle{plain}
\bibliography{literatur,local-refs}


\end{document}
